Competing Bandits: The Perils of Exploration under Competition

### 2018-08-30 paper/competing_bandits.tex
more edits in intro and related work.

\title{Competing Bandits: Exploration and Competition under Duopoly}

@@

Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information). MAB problems have been studied for many decades, \eg see \cite{Bubeck-survey12} for background. We consider a simple and well-understood MAB model with i.i.d. rewards. We focus on a (well-known) distinction between the ``greedy" (exploitation-only) algorithm, "naive" exploration algorithms that separate exploration and exploitation, and more sophisticated algorithms that combine the two.

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system
    \cite{Frazier-ec14,Kremer-JPE14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    \cite{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions
    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.

@@

\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$. We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
\begin{enumerate}
\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation. We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{Shipra-colt12}.
\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

>>>>

%\title{Competing Bandits: Exploration and Competition under Duopoly}

\title{Competing Bandits: The Perils of Exploration under Competition}

@@

Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information). MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background. We consider a simple and well-understood MAB model with i.i.d. rewards \cite{bandits-ucb1}. We focus on a well-known distinction between the ``greedy" (exploitation-only) algorithm, "naive" MAB  algorithms that separate exploration and exploitation, and better MAB algorithms that combine the two. As a token example of ``near-optimal" bandit algorithms, 
we use a classic algorithm called ``Thompson Sampling", see \cite{TS-survey-FTML18} for background.

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system
    \cite{Frazier-ec14,Kremer-JPE14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
pay-per-click auctions
    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm}
%\cite{Amin-auctions-nips13,Jieming-ec18},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.

@@

\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$. We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
\begin{enumerate}
\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation. We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward %\cite{Shipra-colt12}.
\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}
